\documentclass[minion]{homework}
\usepackage{cmacros}
\geometry{top=1in, bottom=0.2in, left=1.2in, right=1.2in}
\doclabel{Math F617: Midterm Exam}
\docdate{March 9, 2012}

\def\ds{\displaystyle}
\def\pts  #1{{\bf[#1 points]}\par}
%\def\worth#1{{\bf[#1 points]}\par}

\begin{document}
{\bf Preamble:} \quad  There is a total of {\bf 56} points on this exam; not every problem is equally weighted.
 Please write out your proofs and your statements of theorems and definitions as clearly as you can.
 In grading, however, some consideration will be made for the exam's time constraints. Good luck!

\begin{aproblems}
\hproblem\pts{14} Define the following.
\begin{subproblems}
  \item A Banach space.
  \item An inner product space.
  \item The dual space of a normed space.
  \item The norm of a linear map $T:X\ra Y$, where $X$ and $Y$
  are normed linear spaces.
  \item A Schauder basis for a Banach space.
  \item An orthonormal basis for a Hilbert space.
  \item A bountiful subset of a metric space.
\end{subproblems}

\hproblem\pts{16} State the following
\begin{subproblems}
	\item H\"older's inequality for $\ell_q$.
	\item Bessel's inequality.
	\item Cauchy-Schwartz inequality.
	\item Parallelogram law.
	\item Baire Category Theorem (any variation is fine).
	\item Banach-Steinhaus Theorem.
	\item Riesz Representation Theorem.
	\item An important characterization of finite dimensional normed spaces.
  % \item Riesz's Lemma concerning distances from subspaces.
\end{subproblems}

\hproblem\pts{16} Let $X$ be a normed space and $Y$ be a Banach space.
Suppose $(T_n)$ is a bounded sequence in $\mathcal{B}(X,Y)$
and that $A$ is a subset of $X$ such that the sequence 
$(T_n)$ converges pointwise on $A$.  Prove that the
sequence also converges pointwise on $\overline{A}$.

\hproblem\pts{10} Let $W$ be a subspace of an inner-product space $X$.
Suppose $x\in X$ and that there exists a closest point $z\in W$
to $x$.  Show that $x-z \in W^\perp$.

\end{aproblems}

\end{document}
