\documentclass[minion,handout]{homework}
\def\ts{\hskip 1.5 em}
\def\tableindent{\hskip 0.5 in}
\def\til{\hbox{\char126}}
\def\heading#1{\localhead{\large#1}}
\def\subheading#1{\localhead{#1}}
\newcommand{\vect}[1]{{\rm\bf #1}}
\newcommand{\vx}{\vect{x}}
\newcommand{\vy}{\vect{y}}
\newcommand{\vv}{\vect{v}}
\newcommand{\vw}{\vect{w}}
\def\Reals{\mathbb{R}}
\newenvironment{clist}%
{\bgroup\parskip 0pt\begin{list}{$\bullet$}{\partopsep 10pt\topsep 0pt\itemsep 0pt}}%
{\end{list}\egroup}%

\doclabel{Math F490: Final Study Ideas}
\docdate{5/1/2015}
\begin{document}
The in-class final will consist of easy computations, statements of definitions, 
and straightforward proofs.  The idea is to encourage you to go back
and review all the things we've learned up to now.  I've listed some study ideas below.
Not everything on the exam is necessarily on this list, and the list certainly too long to have
every topic included in a two-hour exam.

\begin{clist}

\item What is a matrix Lie group?  What is a Lie algebra?  What is a matrix Lie algebra?

\item What is a tangent vector to a matrix Lie group?  What about a tangent vector at the identity?

\item How do you represent reflections using quaternions?

\item How do you represent rotations in $\Reals^4$ and $\Reals^3$ using quaternions?

\item What is the relationship between $SU(2)$ and $SO(4)$?  What is the relationship between
$SU(2)$ and $SO(3)$?  You should knoow the details here.

\item How do you represent complex numbers as matrics?  Quaternions?

\item Why is multiplication by a unit quaterinion an isometry?

\item How are the groups $O(n)$, $SO(n)$, $U(n)$, $SU(n)$, and $Sp(n)$ defined?
What is the common theme connecting these?

\item What are the lie algebras of the groups in the item above? 

\item Know the matrix exponential.  How is it defined?  When is $\exp(A+B)=\exp(A)\exp(B)$?

% \item What is an ideal of a Lie algebra?

\item What is a normal subgroup of a group?  What is the quotient group $G/N$ when
$N$ is a normal subgroup of $G$? What is a simple group?  
%What is a simple Lie algebra?

\item What does it mean for a matrix group to be path connected?  Why is $O(n)$ not
path connected?  

\item What is a maximal torus?  What is the center of a group?  What are the centers of 
the groups $O(n)$, $SO(n)$, $U(n)$, $SU(n)$, and $Sp(n)$?  Why is the center a normal subgroup?

% \item How is $\mathfrak{gl}(n,\mathbb{C})$ related to $\mathfrak{u}(n)$?

\end{clist}
\end{document}
