\documentclass[minion,handout]{homework}
\def\ts{\hskip 1.5 em}
\def\tableindent{\hskip 0.5 in}
\def\til{\hbox{\char126}}
\def\heading#1{\localhead{\large#1}}
\def\subheading#1{\localhead{#1}}
\newcommand{\vect}[1]{{\rm\bf #1}}
\newcommand{\vx}{\vect{x}}
\newcommand{\vy}{\vect{y}}
\newcommand{\vv}{\vect{v}}
\newcommand{\vw}{\vect{w}}
\newcommand{\Reals}{\mathbb{R}}
\newenvironment{clist}%
{\bgroup\parskip 0pt\begin{list}{$\bullet$}{\partopsep 10pt\topsep 0pt\itemsep 0pt}}%
{\end{list}\egroup}%

\doclabel{Math F490: Midterm Topics}
\docdate{4/4/2022}
\begin{document}
The in-class midterm will consist of easy computations, statements of definitions,  and straightforward proofs.  The idea is to encourage you to go back
and review all the things we've learned up to now.  I've listed some study ideas below.
Not everything on the exam is necessarily on this list

\begin{clist}

\item What is a group? A subgroup? A group homomorphisim?

\item What is a normal subgroup of a group?  What is the quotient group $G/N$ when
$N$ is a normal subgroup of $G$? What is a simple group? 

\item Name a theorem that relates reflections and isometries that
preserve the origin in $\Reals^n$.

\item How do you represent rotations in $\Reals^4$ and $\Reals^3$ using quaternions?

\item What is the relationship between $SU(2)$ and $SO(4)$?  What is the relationship between
$SU(2)$ and $SO(3)$?  You should knoow the details here.

\item How do you represent complex numbers as matrics?  Quaternions?

\item Why is multiplication by a unit quaterinion an isometry?

\item How are the groups $O(n)$, $SO(n)$, $U(n)$, $SU(n)$, and $Sp(n)$ defined?
What is the common theme connecting these?


\item How do you represent reflections using quaternions?

\end{clist}
\end{document}
